---
title: "Understanding the Current Data Science and Data Visualization Landscape"
subtitle: "Overview"
author: "Jae Jung"
institute: "Cal Poly Pomona"
date: "`r format(Sys.Date(), '%B, %d, %Y')`"
format: 
  revealjs:
    theme: default
    width: 1600
    height: 900
    transition: slide
    transition-speed: default
    incremental: false
    toc: true
    toc-depth: 2
    slide-number: true
    scrollable: true
    code-overflow: wrap
    code-copy: true
    code-fold: true
    number-sections: true
    number-depth: 5
    embed-resources: true
    css: styles.css
editor: visual
execute: 
  echo: false
  freeze: auto
---

## Various facets of data science

---

### Components
- Domain knowledge
- Statistics/Math
- Programming

---

### Levels
- Descriptive Analytics
- Predictive Analytics
- Prescriptive Analytics


## What is the Role of Visualization?

- Descriptive Analytics
- Predictive Analytics
- Prescriptive Analytics

. . .

- Samples
    - [R Graph Examples](https://r-graph-gallery.com/best-r-chart-examples.html)
    - [Tableau Gallery](https://www.tableau.com/viz-gallery)

## Popular Tools for Data Wrangling and Visualization 

:::{.incremental}
-   R
    -   Base R
    -   Modern R with Tidyverse

-   Python (Pandas/Numpy/Matplotlib/Seaborn)

-   Julia

-   Tableau

-   Power BI

-   Excel
:::

## Pros and cons of the popular tools

:::{.highlight-muted}
-   Modern R with Tidyverse

-   Python (Pandas/Numpy/Matplotlib/Seaborn)

-   Julia

-   Tableau

-   Power BI

-   Excel
:::

## Why Literate Programming (coding)?
:::{.incremental}
-   Reproducibility
-   Transparency
-   Ethics
:::

## Demonstration of data visualization

```{r prep}
#| output: false
library(tidyverse)
library(GGally)

df <- mpg # assigning the data set to a new name
head(df)
theme_set(theme_light()) #set the graphics theme to a light style for this R session.
```

```{r}
#| echo: fenced
#| code-fold: false
#| label: fig-cty-by-cyl-density
#| fig-cap: The impact of cylinder size on the city mileage using density plot
#| fig-cap-location: top
#| output-location: slide

ggplot(df, aes(x = cty)) +
  geom_density(aes(fill=factor(cyl)), alpha=0.8) +
  labs(title = "Density Plot",
       subtitle = "City Mileage Grouped by Number of Cylinders",
       caption = "Source: mpg dataset",
       x= "City Mileage",  
       fill="# Cylinders") 
```

## Interoperability among multiple languages

-   Language agnostic
-   IDE:
    -   RStudio
        -   Collaborating with colleagues who use a different language, using Quarto
    -   Positron
    -   Google Colab
    -   VS Code,
    -   Jupyter notebook
    -   IPython notebook

## Use of AI for data visualization

. . .

::::::{.columns}

::::{.incremental}

:::{.column}
### Pros
-   For experienced users, it helps productivity
-   For beginner, it can be helpful as a tutor
:::

:::{.column}
### Cons
-   It may not give you what you want
-   For beginner, it can be bad for learning and evaluating codes
:::
::::

::::::

## Assignments due next week

-   Spreadsheet Certificate, Due on 2/2
